Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign

### 2022-05-16 analysis/paper/paper_ana_20181122.Rmd
Steve's comments

---
title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
- name: Craig S. Schwartz
  affiliation: NCAR
- name: Stephen W. Nesbitt
  affiliation: Illinois
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign.

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
journal: Atmospheric Research
layout: authoryear,preprint,review,12pt
#layout: final,5p,times,twocolumn,authoryear
address:
- code: UBA
  address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
- code: CIMA
  address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
- code: CNRS
  address: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
- code: SMN
  address: Servicio Meteorológico Nacional de Argentina.
- code: CONICET
  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
- code: INPE
  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.
- code: NCAR
  address: National Center for Atmospheric Research, Boulder, Colorado.
- code: Illinois
  address: Department of Atmospheric Sciences, University of Illinois at Urbana–Champaign, Urbana, Illinois.
bibliography: 
  - paper_ANA.bib 
  - packages.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
keywords: Regional Data Assimilation, Surface Obsevations, Satellite Observations 
numbersections: true
# header-includes: 
#  - \usepackage{gensymb}
#  - \usepackage{subfig}
#  - \usepackage[inline]{showlabels}
#  - \usepackage{chngcntr}
#  - \usepackage{natbib}
#  - \usepackage{lineno}
#  - \linenumbers 
---

@@

Severe weather events cause significant human and material losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

@@

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, the bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given the mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are none studies related to the direct assimilation of radiance observations over South America.

@@

To reach this goal, we conduct several DA experiments for a case study of a gigantic Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.

@@

(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the RELAMPAGO. Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 % of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere. 
The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. First, a spatial data thinning is applied using a 60 km grid following [@singh2016; @jones2013; @lin2017a], where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow [@hu2018].

@@

* Radar observations are used to perform a qualitative and visual assessment of the convection features. The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes. For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.

@@

$$\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}$$

where $M$ is the ensemble size. A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$). If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative. A $sd RCRV > 1$ () indicates that the ensemble is underdispersive and an $sd RCRV < 1$ indicates that the ensemble is overdispersive.

@@

$$\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)$$

where $O$ and $X$ stand for independent observations and the simulations respectively, and N is the sample size.

@@

$$\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} $$

Where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold. Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area. 
The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.

>>>>

---
title: "Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a  Mesoscale Convective System Case During RELAMPAGO campaign"
author:
- name: Paola Belén Corrales
  email: paola.corrales@cima.fcen.uba.ar
  affiliation: UBA,CIMA,CNRS
  footnote: 1
- name: Victoria Galligani
  affiliation: UBA,CIMA,CNRS
- name: Juan Ruiz
  affiliation: UBA,CIMA,CNRS
- name: Luiz Sapucci
  affiliation: INPE
- name: María Eugenia Dillon
  affiliation: SMN,CONICET
- name: Yanina García Skabar
  affiliation: SMN,CONICET,CNRS
- name: Maximiliano Sacco
  affiliation: SMN
- name: Craig S. Schwartz
  affiliation: NCAR
- name: Stephen W. Nesbitt
  affiliation: Illinois
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_book: 
      base_format: rticles::elsevier_article
abstract: |
  In this paper, we evaluate the impact of assimilating high-resolution surface networks and satellite observations using the WRF-GSI-LETKF over central and north eastern Argentina where the surface and upper air observing networks are relatively coarse. We conducted a case study corresponding to a huge mesoscale convective system (MCS) that developed during November 22, 2018. The accumulated precipitation associated with this MCS was quite high, exceeding 200 mm over northern Argentina and Paraguay. The MCS developed during the Intense Observing Period (IOP) of the RELAMPAGO field campaign. We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-member multiphysics ensemble. We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP’s prepBUFR files; AWS combining CONV and dense automatic surface weather station networks, SATWND, combining AWS with satellite-derived winds and RAD, including SATWND; and satellite radiances from different microwave and infrared sensors. We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation. We also ran ensemble forecasts initialized from each experiment and evaluated their skill to predict precipitation. We found that the hourly assimilation of the observations in AWS, SATWND, and RAD helped to improve the precipitation forecast.
journal: Atmospheric Research
layout: authoryear,preprint,review,12pt
#layout: final,5p,times,twocolumn,authoryear
address:
- code: UBA
  address: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
- code: CIMA
  address: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
- code: CNRS
  address: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
- code: SMN
  address: Servicio Meteorológico Nacional de Argentina.
- code: CONICET
  address: CONICET (Consejo Nacional de Investigaciones Científicas y Técnicas).
- code: INPE
  address: National Institute for Space Research, Brazil, Center for Weather Forecasting and Climate Studies.
- code: NCAR
  address: National Center for Atmospheric Research, Boulder, Colorado.
- code: Illinois
  address: Department of Atmospheric Sciences, University of Illinois Urbana–Champaign, Urbana, Illinois.
bibliography: 
  - paper_ANA.bib 
  - packages.bib
csl: elsevier-harvard.csl
footnote:
- code: 1
  text: Corresponding Author
keywords: Regional Data Assimilation, Surface Observations, Satellite Observations 
numbersections: true
header-includes: 
#  - \usepackage{gensymb}
#  - \usepackage{subfig}
#  - \usepackage[inline]{showlabels}
#  - \usepackage{chngcntr}
#  - \usepackage{natbib}
  - \usepackage{lineno}
  - \linenumbers 
---

@@

Severe weather events cause significant human and economic losses around the world. A large number of these phenomena are associated with the occurrence of deep moist convection, including tornadoes, intense wind gusts, extreme precipitation in short time periods, large hail, and lightning. 
Southern South America has one of the highest frequencies in the world of favorable conditions for high-impact meteorological events [@brooks2003] and large hail events [@cecil2012], particularly during austral spring and summer. 
This is also confirmed by observational evidence and high impact weather reports [@matsudo2015; @rasmussen2014]. Recently, the RELAMPAGO-CACTI field campaign [@nesbitt2021] has been conducted to investigate the mechanisms for convective initiation and the occurrence of high-impact weather events associated with deep convection in central Argentina.

@@

While the assimilation of radiance observations into global models is well established [@eyre2020], the direct assimilation of radiance data into regional models, however, still remains a challenge due to the sparse data coverage, bias correction, and the relatively low model tops used for this application. @bao2015 studied the impact of assimilating microwave and infrared radiance data on temperature and humidity forecasts over the western USA and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. More recently, @zhu2019 studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels. @wang2021 studied the impact of assimilating radiances in the high-resolution Copernicus European Regional Reanalysis. They reported that satellite radiance observations had a neutral impact on the analyses of geopotential height in the lower troposphere, while a slightly negative impact on the upper troposphere and the stratosphere. They also observed similar results for 3-h forecasts initialized from the analysis but a positive impact on 12 and 24 -h forecasts. Given these mixed results, there is still room to analyze the utility of assimilating radiance observations in a limited-area DA system over land. Moreover, to the best of our knowledge, there are no studies related to the direct assimilation of radiance observations over South America.

@@

To reach this goal, we conduct several DA experiments for a case study of a large Mesoscale Convective System (MCS) that developed over Southern South America during Nov 22-23, 2018 during the intense observation period (IOP) of the RELAMPAGO field campaign.

@@

(ref:dominio) a) The domain used for the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box), and the locations of Automatic Weather Stations (AWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during RELAMPAGO. Green dots correspond to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8, and purple squares are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations, are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Microwave Humidity Sounder (MHS), and 2 multispectral sensors; the Atmospheric Infrared Sounder (AIRS) and the Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table \@ref(tab:table-rad)). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. In particular, the multispectral sensors provided between 100 and 1000 observations for every scan every 12 hours, contributing 88 % of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere up to the lower stratosphere.

The channels accepted for assimilation and their associated errors were defined taking into account the low model top (50 hPa). The data preprocessing, which is an essential step in the assimilation of radiances, was performed within the GSI system for each sensor specifically. First, a spatial data thinning is applied using a 60 km grid following @singh2016, @jones2013, and @lin2017a, where the observations to be assimilated are chosen based on their distance to the model grid points, the observation quality (based on available data quality information), and the number of available channels (from the same pixel and sensor) that passed the quality control. Also, observations over the sea are preferred to those over land or snow [@hu2018].

@@

* Radar observations are used to perform a qualitative and visual assessment of the convective features. The data comes from 9 radars located in the domain and is provided by the Argentine C-band Doppler dual-polarization weather radar network [@deelia2017] with a temporal frequency of 10 minutes. For this work, only the maximum reflectivity in the column (COLMAX) closest to the analysis time was used.

@@

$$\mathit{sd RCRV} = \sqrt{\frac{M}{M -1}E[(\mathit{RCRV} - \mathit{mean RCRV})^2]}$$
where $M$ is the ensemble size. A consistent system will have no bias ($mean RCRV = 0$) and a standard deviation equal to 1 ($sd RCRV = 1$). If the ensemble has a positive bias, $mean RCRV$ will be positive, on the opposite, if the ensemble has a negative bias, $mean RCRV$ will be negative. A $sd RCRV > 1$ () indicates that the ensemble is underdispersive and an $sd RCRV < 1$ indicates that the ensemble is overdispersive.

@@

$$\mathit{BIAS} = \frac{1}{N}\sum_{i = 1}^{N} (X_i - O_i)$$
where $O$ and $X$ stand for independent observations and the simulations respectively, and N is the sample size.

@@

$$\mathit{FSS} = 1-\frac{\sum_{i=1}^{N} ({P_x}_i-{P_o}_i)^{2}}{\sum_{i=1}^{N} ({P_x}_i)^{2}+\sum_{i=1}^{N} ({P_o}_i)^{2}} $$
where $P_{oi}$ is the fraction of grid points in the $i-th$ sampling area in which the observed accumulated precipitation is greater than a specified threshold. Following @roberts2020, $P_{xi}$ is calculated from the ensemble probability precipitation over the same threshold in each grid point by averaging over the $i-th$ sampling area. 
The FSS was computed from the accumulated precipitation over 6 hr periods by adding the 1-hr accumulated precipitation forecasts over 6 consecutive assimilation cycles.
